\textbf{\sc Towards Deliberative Control in Marine Robotics}

### 2012-02-24 trunk/concepts.tex
heading changes based on Minh and discussion; also noted FP comments
addressed Kanna's comments.

\section{Fundamentals of \eu}
\label{sec:basics}

\paragraph {A brief note on \eu history and notation} The origin of
what we will call \eu in this chapter, is derived from the
\texttt{HSTS} planner \cite{mus94} used on a number of ground-based
demonstrations for space telescope scheduling as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was also the
\texttt{LISP}-based\footnote{This was the first (and to our knowledge
  only) software ever to been written in \texttt{LISP} to be flown in
  space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX} deployment, a C++
version of the planner was implemented which went on to be used for
the \texttt{MAPGEN} \cite{bresina05} system and continues to be used
to date, in the mission-critical uplink process for the
command/control of the Mars Exploration Rovers (\texttt{MER})
mission. A vastly re-factored, higher performance open source version
\cite{europapso} was named \eut and is the subject of this chapter and
used by \rx.  Note however, in using the name \eu, we refer to the
\emph{fundamental ideas as well as the planning paradigm} even if
there exists a specific implementation.

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.35]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eus built-in timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}.  \kcomment{add a figure}

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are a common requirement for \eu
  users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided.
  Instances of classes derived from a resource will induce ordering
  requirements on their tokens in order to ensure that the level of
  the resource remains within specified limits.  \kcomment{This
    definitely needs a visual example and some text esp. since it
    impacts plan structure via scheduling}

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  rules. For instance, when the Shopping Agent must get to a location where a product is
  available before it is able to purchase it: in this case the
  original goal of purchasing the product spawns a subgoal of being at
  a specific location without which the purchase will not occur as
  defined in the domain model.  Looking at how a simple Shopping
  Agent goal would be addresed by \eu helps illustrate how Token
  States support the planning process.  Let's assume the planner
  starts with a partial plan that places the Shopping agent at home at
  time $0$ with no products in possession. Let's also assume that goals are posted for
  the Agent to own a drill by time 10 and to be home by time 20 as depicted in Fig. \ref{fig:europapr1}

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-1.jpg}
  \caption{\small Timelines representing initial state and goals for a Shopping Agent.}
\label{fig:europapr1}
\vskip+0.1cm
\end{figure}

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-merge.jpg}
  \caption{\small Timeline states after satisfying goal to be home by time 20 by merging with initial state}
\label{fig:europapr:merge}
\vskip+0.1cm
\end{figure}

The goal to be home by time 20 is satisfied by merging with the initial state that said that the Agent is home at time 0.
The merge is possible because the tokens are of the same type \texttt{At}, and intersecting the domains for all of their variables
doesn't yield an empty domain (which represents an inconsistency), in this case the token variables are the location, 
which is \texttt{Home} in both cases, and the temporal variables (start, end, duration), which can be intersected without causing 
an inconsistency. After the token representing the goal is merged onto the token representing the initial state, the resulting token
represents not only the fact that the Agent must be \texttt{At(Home)} by time 0 but also \texttt{At(Home)} at time 20, this is reflected in 
the new constraints on the token's start and end variables as depicted in Fig. \ref{fig:europaprmerge}.

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-activate.jpg}
  \caption{\small Timeline states after satisfying goal to own drill by time 10 by activating corresponding token}
\label{fig:europapr:activate}
\vskip+0.1cm
\end{figure}

It is easy to see in Fig. \ref{fig:europapr:activate} that the decision to merge the goal to be \texttt{At(Home)} by time 20 
with the initial state was not a good one since the Agent will have to get away from \texttt{Home} to be able to 
buy the drill which is inconsistent with the constraints that resulted after the merge. As a result that decision will have 
to be retracted at some point and the alternative to activate or merge with a different token will have to be considered.

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-states.jpg}
  \caption{\small Possible transitions for a Token State}
\label{fig:europapr:states}
\vskip+0.1cm
\end{figure}

\item[\textbf{Rules}] In order for a plan to be valid, it must comply
  with all rules pertinent to the relevant application domain. Rules
  govern the internal and external relationships of a token. For
  example, consider a parameterized action
  \texttt{Go(origin,destination)} which describes a Shopping Agent
  traveling from one location (\textit{origin}) to another
  (\textit{destination}). The parameters are instantiated on a token
  as variables whose domain of values is the set of all possible
  locations in a given problem. A rule governing an internal
  relationship among token variables might stipulate that a transition
  must involve a change in location. This can be expressed as a
  constraint on the definition of a predicate of the form:
  \textit{origin != destination}. A further stipulation could be that
  the agent must be located at \textit{origin} before travel can occur
  and the agent must be located at \textit{destination} when travel is
  completed. This is an example of a rule governing an external
  relationship among tokens. It specifies a requirement that tokens of
  the predicate \texttt{At} that represent the location state for the
  Shopping Agent precede and succeed tokens of the predicate
  \texttt{Go} that represents traveling.

  \begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-rules.jpg}
  \caption{\small Master and Slave tokens generated as a result of a Domain Rule}
\label{fig:europapr:rules}
\vskip+0.1cm
\end{figure}

  Fig. \ref{fig:europapr:rules} illustrates the entities and
  relations involved in specifying such a rule on a \texttt{Go}
  action. The token on which the rule applies  (\texttt{Go}) is referred to
 as the \emph{master}. Each \texttt{At} and \texttt{Going} token required
  by the master is referred to as a \emph{slave}. 
  Application of a rule on a token can thus cause slave
  tokens, variables and constraints to be introduced.  The capability
  of a domain rule to cause a slave token to be created is a key
  vehicle through which planning occurs. Semantically, a rule
  imposes a requirement for supporting (slave) tokens to be in the
  plan in order for the master token to be valid.

>>>>

\section{The \eu Planner }
\label{sec:basics}

\paragraph {A brief note on \eu history and notation} The origin of
what we will call \eu in this chapter, is derived from the
\texttt{HSTS} planner \cite{mus94} used on a number of ground-based
demonstrations for space telescope scheduling as noted in Section
\ref{sec:rabeyond}. \texttt{HSTS} was also the
\texttt{LISP}-based\footnote{This was the first (and to our knowledge
  only) software ever to been written in \texttt{LISP} to be flown in
  space.} planner that was flown as part of the Remote Agent
\texttt{RAX} experiment. Concurrent to \texttt{RAX} deployment, a C++
version of the planner was implemented which went on to be used for
the \texttt{MAPGEN} \cite{bresina05} system and continues to be used
to date, in the mission-critical uplink process for the
command/control of the Mars Exploration Rovers (\texttt{MER})
mission. A vastly re-factored, higher performance open source version
\cite{europapso} was named \eut and is the subject of this chapter and
used by \rxe.  Note however, in using the name \eu, we refer to the
\emph{fundamental ideas as well as the planning paradigm} even if
there exists a specific implementation.

@@

\begin{figure}[!htb]
\centering
\includegraphics[scale=0.3]{figs/flexible-timelines.pdf}
\caption{\small Tokens with flexible temporal intervals and parametric
  constraints between tokens. This example shows the triggering of a
  water sampler based on a feature threshold while the vehicle
  Yo-Yo's. The \texttt{Waypoint\_Yo-Yo} token has a flexible duration,
  start \& end times.}
\label{fig:flex-timelines}
\end{figure}

@@

\item \textit{Timelines}: Often objects in a domain must be described
  by exactly one token for every given timepoint in the plan. Any
  instances of a class derived from \eus built-in timeline class will
  induce ordering requirements among its tokens to ensure no temporal
  overlap may occur among them \cite{mus94}. Fig. \ref{fig:europapr1} below and several others
  after that illustrate how timelines are used to represent state in a plan.

\item \textit{Resources}: Metric resources, e.g. the energy of a
  battery or the capacity of a cargo hold, are objects with an
  explicit quantitative state in time and with a circumscribed range
  of changes that can occur to impact that state i.e. produce,
  consume, use, change.  Resources are such a common requirement for \eu
  users that built-in object types (with their corresponding token
  types to denote production, consumption, etc) are provided.
  Instances of classes derived from a resource will induce ordering
  requirements on their tokens in order to ensure that the level of
  the resource remains within specified limits.

@@

\item[\textbf{Token State Model}] In \eu's representation, goals are
  temporally scoped states that need to be achieved, or actions that
  must be performed. As a result, \eu internally represents goals as
  tokens.  Goals can be posted explicitly, for example, when a user
  asks the Shopping Agent to own a specific product by a specific
  time, or implicitly when subgoals are created as a result of domain
  rules. For instance, when the Shopping Agent must get to a location
  where a product is available before it is able to purchase it: in
  this case the original goal of purchasing the product spawns a
  subgoal of being at a specific location without which the purchase
  will not occur as defined in the domain model.  Looking at how a
  simple Shopping Agent goal would be addresed by \eu helps illustrate
  how Token States support the planning process.  Let's assume the
  planner starts with a partial plan that places the Shopping agent at
  home at time $0$ with no products in possession. Let's also assume
  that goals are posted for the Agent to own a drill by time $10$ and
  to be home by time $20$ as shown in Fig. \ref{fig:europapr1}

  \begin{figure} \centering
    \includegraphics[scale=0.4]{figs/europa-pr-1.pdf}
    \caption{\small Timelines representing initial state and goals for a Shopping Agent.}
    \label{fig:europapr1}
  \end{figure}

@@

\begin{figure} 
  \centering
  \includegraphics[scale=0.4]{figs/europa-pr-merge.pdf}
  \caption{\small State of timelines after satisfying a goal to be
    home by time $20$ on merging with initial state.}
  \label{fig:europapr:merge}
\end{figure}

The goal to be home by time $20$ is satisfied by merging with the
initial state that said that the Agent is home at time $0$.  The merge
is possible because the tokens are of the same type \texttt{At}, and
intersecting the domains for all of their variables doesn't yield an
empty domain (which represents an inconsistency), in this case the
token variables are the location, which is \texttt{Home} in both
cases, and the temporal variables (start, end, duration), which can be
intersected without causing an inconsistency. After the token
representing the goal is merged onto the token representing the
initial state, the resulting token represents not only the fact that
the Agent must be \texttt{At(Home)} by time 0 but also
\texttt{At(Home)} at time $20$, this is reflected in the new
constraints on the token's start and end variables as depicted in
Fig. \ref{fig:europapr:merge}.

@@

\begin{figure} \centering
  \includegraphics[scale=0.4]{figs/europa-pr-activate.pdf}
  \caption{\small State of timelines after satisfying a goal to own a
    drill by time $10$ by activating corresponding token.}
\label{fig:europapr:activate}
\end{figure}

As Fig. \ref{fig:europapr:activate} shows the decision to merge the
goal to be \texttt{At(Home)} by time $20$ with the initial state was
not appropriate since the Agent will have to get away from
\texttt{Home} to be able to buy the drill which is inconsistent with
the constraints that resulted after the merge. As a result that
decision will have to be retracted at some point and the alternative
to activate or merge with a different token will have to be
considered.

@@

\begin{figure} \centering
  \includegraphics[scale=0.45]{figs/europa-pr-states.jpg}
  \caption{\small Possible transitions for a Token State}
\label{fig:europapr:states}
\end{figure}

\item[\textbf{Rules}] In order for a plan to be valid, it must comply
  with all rules pertinent to the relevant application domain. Rules
  govern the internal and external relationships of a token. For
  example, consider a parameterized action
  \texttt{Go(origin,destination)} which describes a Shopping Agent
  traveling from one location (\textit{origin}) to another
  (\textit{destination}). The parameters are instantiated on a token
  as variables whose domain of values is the set of all possible
  locations in a given problem. A rule governing an internal
  relationship among token variables might stipulate that a transition
  must involve a change in location. This can be expressed as a
  constraint on the definition of a predicate of the form:
  \textit{origin != destination}. A further stipulation could be that
  the agent must be located at \textit{origin} before travel can occur
  and the agent must be located at \textit{destination} when travel is
  completed. This is an example of a rule governing an external
  relationship among tokens. It specifies a requirement that tokens of
  the predicate \texttt{At} that represent the location state for the
  Shopping Agent precede and succeed tokens of the predicate
  \texttt{Go} that represents traveling. Fig. \ref{fig:europapr:rules}
  illustrates the entities and relations involved in specifying such a
  rule on a \texttt{Go} action. The token on which the rule applies
  (\texttt{Go}) is referred to as the \emph{master}. Each \texttt{At}
  and \texttt{Going} token required by the master is referred to as a
  \emph{slave}.  Application of a rule on a token can thus cause slave
  tokens, variables and constraints to be introduced.  The capability
  of a domain rule to cause a slave token to be created is a key
  vehicle through which planning occurs. Semantically, a rule imposes
  a requirement for supporting (slave) tokens to be in the plan in
  order for the master token to be valid.

  \begin{figure} 
    \centering
    \includegraphics[scale=0.4]{figs/europa-pr-rules.pdf}
    \caption{\small Master and Slave tokens generated as a result of a domain rule.}
    \label{fig:europapr:rules}
  \end{figure}
